---
title: "Crosstable"
author: "Andrés Necochea"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Crosstable}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Crosstable is an extension for the class table that have useful methods to print the table with percents and chisquare expected values.
You can use as any regular table call functions like `apply`, `margin.table` and  `chisq.test`.

## Basic Usage

This function is designed to be used in a way that you can see the results in the console.

```{r}
library(multitabulation)
library(magrittr)

# Generate Random data
gender <- sample(c(1,2), 131, replace=TRUE) %>%
  factor(levels=c(1,2), labels=c("Man", "Woman"))
strata <- sample(c(1,2,3), 131, replace=TRUE) %>%
  factor(levels=c(1,2,3), labels=c("Low", "Middle", "High"))
party <- sample(c(1,2), 131, replace=TRUE) %>%
  factor(levels=c(1,2), labels=c("Right", "Left"))

crosstable(strata, party, gender, stats=c("count", "row", "column", "total"))
```

## Using flextable

You can use `as_flextable` to make a table ready to be printed in a report. You can combine flextable with officer in order to export the table to Ms Office.
```{r, results='asis'}
library(multitabulation)
library(flextable)
library(magrittr)

crosstable(strata, party, gender, stats=c("count", "column")) %>%
  as_flextable
```

## Performing Chi Squared Test

Crosstable is basically an extension of table class, so you can use in almost the same way than a regular table, so you can perform the `chisq.test` 

```{r}
tab <- with(MASS::bacteria, crosstable(y, trt, stats=c("count", "column")))
tabxsq <- chisq.test(tab)
tabxsq
```

Also, you can include the expected values of `chisq.test` in the table report using `add.tables`.

```{r, results='asis'}
add.tables(tab, format="(0.00)", "(expected)" = tabxsq$expected) %>%
  as_flextable
```

## Aditional options

If you prefer the stats on rows, you can use `stats.on.cols=FALSE`

```{r, results='asis'}
crosstable(strata, party, gender, stats=c("count", "column"), stats.on.cols=FALSE) %>%
  as_flextable
```

You can use an existing table and turn it into a `crosstable`. You can use
`col.vars` or `row.vars`

```{r, results='asis'}
crosstable(Titanic, col.vars=c("Sex", "Survived"), stats=c("count", "column")) %>%
  as_flextable
```

You can use `dnn` to set the names of the variables.
```{r, results='asis'}
with(MASS::bacteria, crosstable(y, trt, stats=c("count", "column"), dnn=c("Result", "Treatment"))) %>%
  as_flextable
```

If you use a wheighted survey or data base, you can use the formula mode.

```{r, results='asis'}
crosstable(Freq ~ Type + Cont, stats=c("count", "row"), data=MASS::housing) %>%
  as_flextable
```
